---
title: Make beautiful graphs with R
author: David Odongo
date: '2019-06-12'
slug: make-beautiful-graphs-with-r
categories:
  - R
  - Data Visualisation
tags:
  - ggplot2khhk
  - bbplot
  - dplyr
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = F, message = F)
```

 Greetings to you my friend, today i would like to talk about how to make charts in R. We shall use the [`ggplot2`](https://ggplot2.tidyverse.org/) package which follows the rules of grammar of graphics.

> A grammar of graphics is a tool that enables us to concisely describe the components of a graphic. Such a grammar allows us to move beyond named graphics (e.g., the ``scatterplot'') and gain insight into the deep structure that underlies statistical graphics

We shall first start by loading the libraries we are going to use.

```{r}
#Load the libraries
library(dplyr) # data wrangling and manipulation of the dataset
library(ggplot2) # using the grammar of graphics to create charts
library(bbplot) #using the beautiful standards bbc uses to develop it's graphics
library(knitr) # pleasant looking tables (not its only use though)
library(gameofthrones) #nice colour palletes based on Game of thrones series
```

We are going to use the diamonds dataset from the `ggplot2` package. We shall assign it to "raw_data"
```{r }
#load the dataset
raw_data <- ggplot2::diamonds

#inspect the dataset
dim(raw_data)
```

This dataset has `r nrow(raw_data) ` rows and `r ncol(raw_data)` columns. 
```{r}
#look at the variables in the dataset
dplyr::glimpse(raw_data)

#look  at the first 5 rows of the dataset
utils::head(raw_data) %>% knitr::kable()

#look at the last 5 rows to make sure the data is consistent
utils::tail(raw_data) %>% knitr::kable()
```


## Visualise the data

Let's now visualise the data to get an easier interpreation of the dataset

```{r}
# Histogram of price
raw_data %>%  ggplot2::ggplot(aes(x = price, fill = cut)) +
  geom_histogram()+
  ggtitle("Price distribution by type of Cut")
```

As you can see, R has really beautiful graphics, but we can make this chart look even better by using the `bbplot` package and also choosing our colour pallete using the `gameofthrones` package

```{r}
# Histogram of price
raw_data %>%  ggplot2::ggplot(aes(x = price, fill = cut)) +
  ggplot2::geom_histogram(colour = 'white')+
  ggplot2::ggtitle("Price distribution by type of Cut") +
  gameofthrones::scale_fill_got_d(option = "Daenerys") +
  ggplot2::geom_hline(yintercept = 0, size = 1, colour="#333333") +
  ggplot2::scale_x_continuous(breaks = c(5000, 10000, 15000)) +
  bbplot::bbc_style()
```

